---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
```
Tibble in long data format that contains the maximum temperature in Scotland over 106 years (1910 to 2015)
```{r}
temp_sco <- CodeClanData::temp_df
```


```{r}
temp_sco %>%
  glimpse()
```



Mean and StD temperature of each month
```{r}
temp_sco %>% 
  group_by(month) %>% 
  summarise(avg_temp = mean(max_temp, na.rm = TRUE),
            med_temp = median(max_temp, na.rm = TRUE),
            std_temp = sd(max_temp, na.rm = TRUE))
```


Mean and StD temperature of each year
```{r}
temp_sco %>% 
  group_by(year) %>% 
  summarise(avg_annual_temp = mean(max_temp, na.rm = TRUE),
            std_annual_temp = sd(max_temp, na.rm = TRUE))
```

Overall Mean and StD temperature
```{r}
temp_sco %>%
  summarise(avg_temp = mean(max_temp, na.rm = TRUE),
            med_temp = median(max_temp, na.rm = TRUE),
            std_temp = sd(max_temp, na.rm = TRUE))
```



Find the year with the max temperature recorded
```{r}
temp_sco %>% 
  group_by(year) %>% 
  summarise(max_temp = max(max_temp, na.rm = TRUE)) %>% 
  slice_max(max_temp)
```


Arrange the maximum recorded temperatures by years and month
```{r}
temp_sco %>% 
  group_by(year, month) %>% 
  summarise(max_temp = max(max_temp, na.rm = TRUE)) %>% 
  slice_max(max_temp) %>% 
  arrange(desc(max_temp))
```


Plot the variability of max temperature for each month (years from 1910 to 2015) - ver. 1
```{r}
temp_sco %>% 
ggplot(mapping = aes(x = year, y = max_temp)) +
  geom_line() +
  facet_wrap(~ month)
  
```
Plot the variability of max temperature for each month (years from 1910 to 2015) - ver. 2
```{r}
temp_sco %>% 
ggplot(mapping = aes(x = year, y = max_temp, colour = month)) +
  geom_line() +
  facet_wrap(~ month, ncol = 12)
  
```

Plot the variability of max temperature for each month (years from 1910 to 2015) - ver. 3
```{r}
temp_sco %>% 
ggplot(mapping = aes(x = year, y = max_temp, colour = month)) +
  geom_line() 
  
```



Plot the variability of max temperature for each month (years from 1910 to 2015)
```{r}
temp_sco %>% 
  ggplot(mapping = aes(x = month, y = max_temp)) +
  geom_point(colour = "blue")
  
```


```{r}
temp_sco %>% 
  filter(month == "jan") %>% 
  ggplot(mapping = aes(x = year, y = max_temp)) +
  geom_line() +
  geom_point() +
  geom_smooth(method = "lm", show.legend = FALSE) 
```


```{r}
temp_sco %>% 
  filter(month == "apr") %>% 
  ggplot(mapping = aes(x = max_temp)) +
  # geom_histogram(binwidth = 0.5, col = "white")
   geom_histogram(bins = 30, col = "white")
```

Plot the max temperature distribution for each month (years from 1910 to 2015)
```{r}
temp_sco %>% 
  ggplot(mapping = aes(x = max_temp, fill = year)) +
  geom_histogram(binwidth = 1, col = "white")+
  # geom_histogram(bin = 25, col = "white")+
  facet_wrap(~ month, nrow = 4)

```

Plot trend of the max temp vs year (1910 to 2015)
```{r}
temp_sco %>% 
  group_by(year, month) %>% 
  summarise(max_temp = max(max_temp, na.rm = TRUE)) %>% 
  slice_max(max_temp, with_ties = FALSE) %>% 
  ggplot(mapping = aes(x = year, y = max_temp)) +
  geom_line() +
  geom_point() 
```


