---
title: "R Notebook"
output: html_notebook
---

# Preamble

## Call libraries
```{r}
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(ggthemes)
```

# Number of cancer cases - Data Wrangling

## View cancer incidence data 


**Incidence by Health Board**

``https://www.opendata.nhs.scot/dataset/c2c59eb1-3aff-48d2-9e9c-60ca8605431d/resource/3aef16b7-8af6-4ce0-a90b-8a29d6870014/download/opendata_inc9620_hb.csv``

**Note:** I had to use the following command and arguments to read the csv file:
```{r}
incidence <- read_delim(here("raw_data/opendata_inc9620_hb.csv"), delim = "\t") %>% 
  clean_names()


glimpse(incidence)

```


### Check incidence data for NA values
```{r}

incidence  %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```
No missing values!!!


## View 5 year summary cancer incidence data 


**5 Year Summary of Incidence by Health Board**

``https://www.opendata.nhs.scot/dataset/c2c59eb1-3aff-48d2-9e9c-60ca8605431d/resource/e8d33b2b-1fb2-4d59-ad21-20fa2f76d9d5/download/opendata_inc1620comb_hb.csv``

**Note:** I had to use the following command and arguments to read the csv file:
```{r}
incidence_5_year_summary <- read_delim(here("raw_data/opendata_inc1620comb_hb.csv"), delim = "\t") %>% 
  clean_names()


glimpse(incidence_5_year_summary)

```


### Check 5 year summary incidence data for NA values
```{r}

incidence_5_year_summary  %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```
No missing values!!!

### What are the distinct cancer values in the incidence data ?

Check the diagnosed cancer site
```{r}
incidence  %>%
  distinct(cancer_site)
```


### What are the distinct cancer values in the 5 year summary incidence data ?

Check the diagnosed cancer site
```{r}
incidence_5_year_summary  %>%
  distinct(cancer_site)
```

## View health board data

**Health Board 2014 - Health Board 2019**

``https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/652ff726-e676-4a20-abda-435b98dd7bdc/download/hb14_hb19.csv``

```{r}
health_board_labels <- read_csv(here("raw_data/geography_codes_and_labels_hb2014_01042019.csv")) %>% 
  clean_names()

glimpse(health_board_labels)
```

### View all HB values
```{r}
health_board_labels %>% 
  distinct(hb, hb_name, hb_date_enacted)
```

### Remove junk
```{r}
health_board_labels <- health_board_labels %>% 
  select(hb, hb_name)
  

glimpse(health_board_labels)
```

## View Custom Grouped Geographies

**Grouped geographies used by ISD Scotland for bespoke analysis or to aid the availability of data which usually would have to be suppressed**

``https://www.opendata.nhs.scot/dataset/65402d20-f0f1-4cee-a4f9-a960ca560444/resource/8f0c3067-7b10-44c6-af36-37f87a9e6efa/download/grouped-geography.csv``

```{r}
grouped_geography_labels <- read_csv(here("raw_data/grouped-geography.csv")) %>% 
  clean_names()

glimpse(grouped_geography_labels)
```

### Remove junk

```{r}
grouped_geography_labels <- grouped_geography_labels %>% 
  select(grouped_geography, grouped_geography_name)
  

glimpse(grouped_geography_labels)
```


# Join tables 

## Join 5 year summary incidence & health board & custom grouped geographies
```{r}
incidence_5_year_summary <- incidence_5_year_summary %>%
  left_join(health_board_labels, "hb")

incidence_5_year_summary <- incidence_5_year_summary %>%
  left_join(grouped_geography_labels, by = c("hb" = "grouped_geography"))

glimpse(incidence_5_year_summary)
```


### Concatenate ``hb_name`` and ``grouped_geography_name`` in a single column called ``hb_name``
```{r}
incidence_5_year_summary <- incidence_5_year_summary %>%
  mutate(hb_name = if_else(!is.na(hb_name), hb_name, grouped_geography_name)) %>%
  select(-grouped_geography_name)

glimpse(incidence_5_year_summary)
```

```{r}

incidence_5_year_summary  %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```


## Join incidence & health board
```{r}
incidence <- incidence %>%
  left_join(health_board_labels, "hb")

incidence <- incidence %>%
  left_join(grouped_geography_labels, by = c("hb" = "grouped_geography"))


glimpse(incidence)
```

### Concatenate ``hb_name`` and ``grouped_geography_name`` in a single column called ``hb_name``
```{r}
incidence <- incidence %>%
  mutate(hb_name = if_else(!is.na(hb_name), hb_name, grouped_geography_name)) %>%
  select(-grouped_geography_name)

glimpse(incidence_5_year_summary)
```

```{r}

incidence  %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```

**Consider Saving in the clean_data folder your incidence data**



#  Data Investigation

## Where is NHS Borders standing compared to other NHBs?

```{r}
incidence_5_year_summary %>% 
  filter(cancer_site == "All cancer types", sex == "All") %>% 
  mutate(label_text = ifelse(hb_name == "NHS Borders", "The NHS Borders has 4065 incidents", NA)) %>% 
  select(hb_name, incidences_all_ages, label_text) %>% 
  ggplot() +
  aes(x = reorder(hb_name, incidences_all_ages), y = incidences_all_ages, fill = reorder(hb_name, incidences_all_ages) == "NHS Borders") +
  geom_col() +
  scale_fill_manual(values = c('steelblue', 'red'), guide = "none") +
  coord_flip() +
  geom_text(aes(label = label_text), fontface = "bold",                  
           color = "red", size = 3, hjust = -0.05) 
```

## Incidence trend

Trend plot of all cancer types
```{r}
incidence %>% 
  filter(cancer_site == "All cancer types", hb_name == "NHS Borders") %>% 
  select(cancer_site, sex, year, incidences_all_ages) %>% 
  ggplot() +
  aes(x = year, y = incidences_all_ages, group = sex, colour = sex) +
  geom_line() +
  geom_point() +
  scale_colour_brewer(palette = "Dark2")
  
  
```

## 5 Year Summary 

Top 5 most common cancer types across all ages for 2016-2020:
```{r}
incidence_5_year_summary %>%
  filter(sex == "All", cancer_site != "All cancer types", hb_name == "NHS Borders") %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_max(incidences_all_ages, n = 5, with_ties = TRUE)
  # arrange(desc(incidences_all_ages))
```

```{r}
incidence_5_year_summary %>%
  filter(sex == "All", cancer_site != "All cancer types", hb_name == "NHS Borders") %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_max(incidences_all_ages, n = 5, with_ties = TRUE) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_all_ages), y = incidences_all_ages) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = incidences_all_ages), fontface = "bold", color = "white", size = 3, hjust = 1.25) +
  coord_flip()
```


<br>
Top 5 most common cancer types for males across all ages for 2016-2020:
```{r}
incidence_5_year_summary %>%
  filter(cancer_site != "All cancer types", sex == "Male", hb_name == "NHS Borders") %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_max(incidences_all_ages, n = 5, with_ties = TRUE)
  # arrange(desc(incidences_all_ages))
```


```{r}
incidence_5_year_summary %>%
  filter(cancer_site != "All cancer types", sex == "Male", hb_name == "NHS Borders") %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_max(incidences_all_ages, n = 5, with_ties = TRUE) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_all_ages), y = incidences_all_ages) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = incidences_all_ages), fontface = "bold", color = "white", size = 3, hjust = 1.5) +
  coord_flip()
```



<br>
Top 5 most common cancer types for females across all ages for 2016-2020:
```{r}
incidence_5_year_summary %>%
  filter(cancer_site != "All cancer types", sex == "Female", hb_name == "NHS Borders") %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_max(incidences_all_ages, n = 5, with_ties = TRUE)
  # arrange(desc(incidences_all_ages))
```

```{r}
incidence_5_year_summary %>%
  filter(cancer_site != "All cancer types", sex == "Female", hb_name == "NHS Borders") %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_max(incidences_all_ages, n = 5, with_ties = TRUE) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_all_ages), y = incidences_all_ages) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = incidences_all_ages), fontface = "bold", color = "white", size = 3, hjust = 1.5) +
  coord_flip()
```


<br>
Most affected age group from cancer for 2016-2020:
```{r}
incidence_5_year_summary_longer <- incidence_5_year_summary %>%
  filter(cancer_site == "All cancer types", sex == "All", hb_name == "NHS Borders") %>% 
  select(cancer_site, starts_with("incidences_age")) %>%
  pivot_longer(-cancer_site, names_to = "incidences_age", values_to = "age_group_count") %>%
  separate(incidences_age, c("extra_column","age_group"), sep ="incidences_age_*") %>%
  select(-extra_column) 

incidence_5_year_summary_longer <- incidence_5_year_summary_longer %>% 
  mutate(age_group = str_replace(age_group, "to", "-")) %>% 
  mutate(age_group = str_replace(age_group, "under", "<")) %>% 
  mutate(age_group = str_replace(age_group, "and_over", "+")) %>% 
  mutate(age_group = factor(
    age_group, 
    levels = c("<5", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", "35-39",
               "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74",
               "75-79", "80-84", "85+"
               )
           )
        )

incidence_5_year_summary_longer %>% 
  group_by(age_group) %>% 
  summarise(max_count = max(age_group_count)) %>% 
  arrange(desc(max_count))
```


```{r}
incidence_5_year_summary_longer %>% 
  group_by(age_group) %>% 
  summarise(max_count = max(age_group_count)) %>%
  ggplot() +
  aes(x = age_group, y = max_count) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = max_count), fontface = "bold", color = "black", size = 3, vjust = -0.5) +
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5))

```





# Misc



<br>
Trend plot of Non-melanoma skin cancer (most common cancer types across all ages for 2016-2020)
```{r}
incidence %>% 
  filter(cancer_site == "Non-melanoma skin cancer", hb_name == "NHS Borders") %>% 
  select(cancer_site, sex, year, incidences_all_ages) %>% 
  ggplot() +
  aes(x = year, y = incidences_all_ages, group = sex, colour = sex) +
  geom_line() +
  geom_point() +
  scale_colour_brewer(palette = "Dark2")
  
  
```



<br>
Top 5 least common cancer types across all ages for 2016-2020:
```{r}
incidence_5_year_summary %>%
  filter(sex == "All", cancer_site != "All cancer types", hb_name == "NHS Borders") %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_min(incidences_all_ages, n = 5, with_ties = TRUE)
  # arrange(desc(incidences_all_ages))
```



```{r}
incidence_5_year_summary %>%
  filter(sex == "All", cancer_site != "All cancer types", hb_name == "NHS Borders") %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_min(incidences_all_ages, n = 5, with_ties = TRUE) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_all_ages), y = incidences_all_ages) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = incidences_all_ages), fontface = "bold", color = "white", size = 3, hjust = 1.5)+
  coord_flip()
```




















