---
title: "R Notebook"
output: html_notebook
---

# Preamble

## Call libraries
```{r}
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(ggthemes)
```

# Number of cancer cases - Data Wrangling

## View cancer incidence data 


**Incidence by Health Board**

``https://www.opendata.nhs.scot/dataset/c2c59eb1-3aff-48d2-9e9c-60ca8605431d/resource/3aef16b7-8af6-4ce0-a90b-8a29d6870014/download/opendata_inc9620_hb.csv``

**Note:** I had to use the following command and arguments to read the csv file:
```{r}
incidence <- read_delim(here("raw_data/opendata_inc9620_hb.csv"), delim = "\t") 


glimpse(incidence)

```


### Check incidence data for NA values
```{r}

incidence  %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```
No missing values!!!


## View 5 year summary cancer incidence data 


**5 Year Summary of Incidence by Health Board**

``https://www.opendata.nhs.scot/dataset/c2c59eb1-3aff-48d2-9e9c-60ca8605431d/resource/e8d33b2b-1fb2-4d59-ad21-20fa2f76d9d5/download/opendata_inc1620comb_hb.csv``

**Note:** I had to use the following command and arguments to read the csv file:
```{r}
incidence_5_year_summary <- read_delim(here("raw_data/opendata_inc1620comb_hb.csv"), delim = "\t") 


glimpse(incidence_5_year_summary)

```


### Check 5 year summary incidence data for NA values
```{r}

incidence_5_year_summary  %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```
No missing values!!!

### what are the distinct cancer values in the incidence data ?

Check the diagnosed cancer site
```{r}
incidence  %>%
  distinct(CancerSite)
```


### what are the distinct cancer values in the 5 year summary incidence data ?

Check the diagnosed cancer site
```{r}
incidence_5_year_summary  %>%
  distinct(CancerSite)
```

## View health board data

**Health Board 2014 - Health Board 2019**

``https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/652ff726-e676-4a20-abda-435b98dd7bdc/download/hb14_hb19.csv``

```{r}
health_board_labels <- read_csv(here("raw_data/geography_codes_and_labels_hb2014_01042019.csv")) 

glimpse(health_board_labels)
```

### View all HB values
```{r}
health_board_labels %>% 
  distinct(HB, HBName, HBDateEnacted)
```

### Remove junk
```{r}
health_board_labels <- health_board_labels %>% 
  select(HB, HBName)
  

glimpse(health_board_labels)
```


# Join tables 

## Join 5 year summary incidence & health board
```{r}
incidence_5_year_summary <- incidence_5_year_summary %>%
  left_join(health_board_labels, "HB")

glimpse(incidence_5_year_summary)
```


## Join incidence & health board
```{r}
incidence <- incidence %>%
  left_join(health_board_labels, "HB")

glimpse(incidence)
```


#  Tidy up in-scope incidence tables

## Incidence table
```{r}
incidence <- incidence %>%
  filter(HBName == "NHS Borders") %>% 
  clean_names()

glimpse(incidence)
```

## 5 Year Summary Incidence table
```{r}
incidence_5_year_summary <- incidence_5_year_summary %>%
  filter(HBName == "NHS Borders") %>% 
  clean_names()

glimpse(incidence_5_year_summary)
```

**Consider Saving in the clean_data folder your incidence data**

Ratio of males vs females:
```{r}
incidence_5_year_summary %>%
  group_by(sex) %>% 
  summarise(cancer_counts = n())
```


<br>
Bike hires gender distribution:
```{r}
nyc_bikes_df %>%
  index_by(year) %>%
  group_by(gender) %>% 
  summarise(bike_hire_counts = n(), bike_hiring_percentage = 100*bike_hire_counts/nrow(nyc_bikes_df))
```





















































