---
title: "NHS Borders Cancer Analysis"
author: "John Hios"
date: "05/05/2022"
output:
  html_document:
    keep_md: yes
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    theme: cerulean #flatly #journal #cosmo
  pdf_document: default
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center')
```


# Load Data


Load libraries:
```{r, message=FALSE, warning=FALSE, echo = TRUE}
library(tidyverse)
library(janitor)
library(here)
library(ggthemes)
```

Run clean data R script:
```{r, message=FALSE, warning=FALSE, echo = TRUE}
source(here("scripts/clean_nhs_borders_data.R"))
```

# Visualisations

## NHS Borders Ranking Against Other NHBs

```{r}
incidence_5_year_summary %>% 
  filter(cancer_site == "All cancer types", sex == "All") %>% 
  mutate(label_text = ifelse(hb_name == "NHS Borders", 
                             "The NHS Borders has 4065 incidents", NA)) %>% 
  select(hb_name, incidences_all_ages, label_text) %>% 
  ggplot() +
  aes(
    x = reorder(hb_name, incidences_all_ages), 
    y = incidences_all_ages, 
    fill = reorder(hb_name, incidences_all_ages) == "NHS Borders"
  ) +
  geom_col() +
  scale_fill_manual(values = c('steelblue', 'red'), guide = "none") +
  coord_flip() +
  geom_text(aes(label = label_text), fontface = "bold",                  
           color = "red", size = 3, hjust = -0.05) +
  ggtitle(
    "Cancer Incidence Across NHS Boards In Scotland",
    subtitle =  "(All cancer types* ; Years 2016-2020)"
  ) +
  labs(x = "Health Board \n ", 
       y = "\n Number of Cases", 
       caption = "\n * excluding non-melanoma skin cancers (ICD-10 C00-C97 excl C44)"
  ) 
```


<br>
```{r}
incidence_5_year_summary %>% 
  filter(cancer_site == "All cancer types", sex == "All") %>% 
  select(hb_name, easr) %>% 
  ggplot() +
  aes(
    x = reorder(hb_name, easr), 
    y = easr, 
    fill = reorder(hb_name, easr) == "NHS Borders"
  ) +
  geom_col() +
  scale_fill_manual(values = c('steelblue', 'red'), guide = "none") +
  coord_flip() +
  ggtitle(
    "Cancer Incidence Across NHS Boards In Scotland",
    subtitle =  "(All cancers*; Rate per 100k person-years at risk; Years 2016-2020)"
  ) +
  labs(x = "Health Board \n ", 
       y = "\n European Age Standardised Rate (EASR)", 
       caption = "\n * excluding non-melanoma skin cancers (ICD-10 C00-C97 excl C44)"
  ) 
```


## Cancer Incident Trend

```{r}
incidence %>% 
  filter(cancer_site == "All cancer types", hb_name == "NHS Borders", ) %>% 
  select(cancer_site, sex, year, incidences_all_ages) %>% 
  ggplot() +
  aes(x = year, y = incidences_all_ages, group = sex, colour = sex) +
  geom_line(size = 0.75) +
  geom_point(size = 1.5) +
  scale_colour_brewer(palette = "Dark2") +
  ggtitle(
    "NHS Boarders Cancer Incidence Trend Plot",
    subtitle =  "(All cancer types*; Years 1996-2020)"
  ) +
  labs(x = "\nYear", 
       y = "Number of Cases \n",  
       caption = "\n * excluding non-melanoma skin cancers (ICD-10 C00-C97 excl C44)",
       color = "Gender"
  ) 
```


## Overview Of 5-Year Summary Data 


```{r}
incidence_5_year_summary %>%
  filter(
    sex == "All", 
    cancer_site != "All cancer types", 
    hb_name == "NHS Borders"
  ) %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_max(incidences_all_ages, n = 5, with_ties = TRUE) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_all_ages), y = incidences_all_ages) +
  geom_col(fill = "steelblue") +
  geom_text(
    aes(label = incidences_all_ages), 
    fontface = "bold", 
    color = "white", 
    size = 3, 
    hjust = 1.25
  ) +
  coord_flip() +
  ggtitle(
    "Top 5 Cancer Types Across All Ages",
    subtitle =  "(Years 2016-2020)"
  ) +
  labs(x = "Year \n", y = "\n Number of Cases") 
```



<br>
```{r}
incidence_5_year_summary %>%
  filter(
    cancer_site != "All cancer types", 
    sex == "Male", 
    hb_name == "NHS Borders"
  ) %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_max(incidences_all_ages, n = 5, with_ties = TRUE) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_all_ages), y = incidences_all_ages) +
  geom_col(fill = "steelblue") +
  geom_text(
    aes(label = incidences_all_ages), 
    fontface = "bold", 
    color = "white", 
    size = 3, 
    hjust = 1.5
  ) +
  coord_flip() +
  ggtitle(
    "Top 5 Cancer Types For Males Across All Ages",
    subtitle =  "(Years 2016-2020)"
  ) +
  labs(x = "Year", y = "\n Number of Cases") +
  ylim(c(0, 1000))
```



<br>
```{r}
incidence_5_year_summary %>%
  filter(
    cancer_site != "All cancer types", 
    sex == "Female", 
    hb_name == "NHS Borders"
  ) %>% 
  select(cancer_site, incidences_all_ages) %>% 
  slice_max(incidences_all_ages, n = 5, with_ties = TRUE) %>% 
  ggplot() +
  aes(x = reorder(cancer_site, incidences_all_ages), y = incidences_all_ages) +
  geom_col(fill = "steelblue") +
  geom_text(
    aes(label = incidences_all_ages), 
    fontface = "bold", 
    color = "white", 
    size = 3, 
    hjust = 1.5
  ) +
  coord_flip() +
  ggtitle(
    "Top 5 Cancer Types For Females Across All Ages",
    subtitle =  "(Years 2016-2020)"
  ) +
  labs(x = "Year \n", y = "\n Number of Cases")
```



<br>
```{r, message=FALSE, warning=FALSE, echo = FALSE}
# Change tibble format to long
incidence_5_year_summary_longer <- incidence_5_year_summary %>%
  filter(
    cancer_site == "All cancer types", 
    hb_name == "NHS Borders"
  ) %>% 
  select(cancer_site, sex, starts_with("incidences_age")) %>%
  pivot_longer(
    -c(cancer_site, sex), 
    names_to = "incidences_age", 
    values_to = "cancer_incidences"
  ) %>%
  separate(
    incidences_age, 
    c("extra_column","age_group"), 
    sep ="incidences_age_*"
  ) %>%
  select(-extra_column) 


# Get age group values in presentable format and place in correct order
incidence_5_year_summary_longer <- incidence_5_year_summary_longer %>% 
  mutate(age_group = str_replace(age_group, "to", "-")) %>% 
  mutate(age_group = str_replace(age_group, "under", "<")) %>% 
  mutate(age_group = str_replace(age_group, "and_over", "+")) %>% 
  mutate(
    age_group = factor(
      age_group, 
      levels = c("<5", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
                 "35-39", "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", 
                 "70-74","75-79", "80-84", "85+" )
    )
  )

```



<br>
```{r}
incidence_5_year_summary_longer %>% 
  filter(sex == "All") %>% 
  group_by(age_group) %>% 
  select(age_group, sex, cancer_incidences) %>% 
  ggplot() +
  aes(x = age_group, y = cancer_incidences) +
  geom_col(fill = "steelblue") +
  geom_text(
    aes(label = cancer_incidences), 
    fontface = "bold", 
    color = "black", 
    size = 3, 
    vjust = -0.5
  ) +
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5)) +
  ggtitle(
    "Cancer Incidents By Age",
    subtitle =  "(All cancer types*; Years 1996-2020)"
  ) +
  labs(x = "\n Age Group", 
       y = "Number of Cases \n",
       caption = "\n * excluding non-melanoma skin cancers (ICD-10 C00-C97 excl C44)"
  )
```



```{r}
incidence_5_year_summary_longer %>% 
  filter(sex != "All") %>% 
  group_by(age_group) %>% 
  select(age_group, sex, cancer_incidences) %>% 
  ggplot() +
  aes(x = age_group, y = cancer_incidences, fill = sex) +
  geom_col(position = "dodge") +
  scale_fill_manual(values=c("blue", "steelblue")) +
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5)) +
  ggtitle(
    "Cancer Incidents By Age and Gender",
    subtitle =  "(All cancer types*; Years 1996-2020)"
  ) +
  labs(x = "\n Age Group", 
       y = "Number of Cases \n", 
       fill = "Gender",
       caption = "\n * excluding non-melanoma skin cancers (ICD-10 C00-C97 excl C44)"
  )
```


# Insights

- The NHS Borders are on the lower end of cancer cases in Scotland (years 2016-2020).
- The overall number of cancer incidents is steadily increasing with time (years 1996-2020). 
- The cancer incidents between males and females are comparable.
- The drop in the cancer incidents observed in 2019 and 2020 is mainly driven by the drop in male cancer cases.
- Skin type cancers are the most common for all genders in 2016-2020. They account for 60% of the top 5 cancer incidents.
- Prostate ranks No.3 among the most common cancers for males (years 2016-2020).
- Breast cancer ranks No.2 among the most common cancers for females (years 2016-2020).
- The most affected age group from cancer is between ``70-74``, followed by the ``65-69`` and ``75-79`` age groups. These 3 age groups account for approx. 50% of the cancer cases in the NHS Borders (years 2016-2020).
- Cancer incidences are higher for females in the ``25-59`` range and males in the ``60-79`` range (different cancer types affecting males/females)
