---
title: "Clustering homework"
author: "John Hios"
date: "19/05/2022"
output:
  html_document:
    keep_md: yes
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    theme: cerulean #flatly #journal #cosmo
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


* You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings. 

**Ans**

Call Libraries
```{r}
library(janitor)
library(here)
library(fastDummies)
library(broom)
library(tidyverse)
library(factoextra)
```


Load data
```{r}
customers <- read_csv(here("data/mall_customers.csv")) %>% 
  clean_names()

glimpse(customers)
```


Check for missing values
```{r}
customers  %>% 
  summarise(across(everything(), .fns = ~sum(is.na(.x))))
```

View mean and std
```{r}
customers %>%
  group_by(gender) %>%
  summarise(
    mean_age = mean(age),
    sd_age = sd(age),
    mean_income = mean(annual_income_k),
    sd_income = sd(annual_income_k),
    mean_spending_score = mean(spending_score_1_100),
    sd_spending_score = sd(spending_score_1_100)
  )
```






* We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers. 

**Ans**

Create scatterplot of ``annual_income_k`` vs ``spending_score_1_100``

```{r}
customers %>%
  ggplot() +
  aes(x = annual_income_k, y = spending_score_1_100, color = gender) +
  geom_point()
```

It seems that there are 5 candidate clusters in the scatterplot


* Perform k-means clustering and chose a value of k.

**Ans**

```{r}
customers_subset <- customers %>% 
  select(annual_income_k, spending_score_1_100)
```

```{r}
set.seed(1234)

clustered_customers <- kmeans(customers_subset, 
                            centers = 5, 
                            nstart = 25)
clustered_customers
```


```{r}
tidy(clustered_customers, 
     col.names = colnames(customers_subset))
```


```{r}
augment(clustered_customers, customers_subset)
```

Clustered Metrics
```{r}
glance(clustered_customers)
```



Gathering cluster info
```{r}
# Set min & max number of clusters want to look at 
max_k <- 8 

k_clusters <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(customers_subset, .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, customers_subset)
  )

k_clusters
```

Choosing Number of clusters

```{r}
clusterings <- k_clusters %>%
  unnest(glanced)

clusterings
```



1. Elbow method
```{r}
fviz_nbclust(customers_subset, 
             kmeans, 
             method = "wss", 
             nstart = 25)
```


2. Silhouette coefficient
```{r}
fviz_nbclust(customers_subset, kmeans, method = "silhouette", nstart = 25)
```


3. Gap statistic
```{r}
fviz_nbclust(customers_subset, kmeans, method = "gap_stat", nstart = 50) #would put nstart=25 if had more computing power
```


Methods #1 and #2 indicate k=5 clusters is sufficient



```{r}
clusterings %>% 
  unnest(augmented) %>%
  filter(k == 5) %>%
  group_by(.cluster) %>%
  summarise(mean(annual_income_k), mean(spending_score_1_100))
```



* Visualise the clustering for your chosen value of k.

**Ans**

```{r}
clusterings %>% 
  unnest(cols = c(augmented)) %>% 
  filter(k == 5) %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point(aes(color = .cluster))
```




* Do you think the clustering seems a good fit for this data?

**Ans**

It seems a reasonable fit, which is in line with human intuition based on the scatterplot created during the exploratory analysis stage of  the data.

* Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the 'Segmentation & clustering intro' lesson). 

**Ans**





