---
title: "R Notebook"
output: html_notebook
---

<hr>

# MVP

<br>

1. Load the `housing_prices.csv` data set and undertake an initial exploration of the data. You will find details on the data set on the relevant [Kaggle page](https://www.kaggle.com/camnugent/california-housing-prices)

<br>
**Ans**

Call libraries
```{r}
library(tidyverse)
library(here)
library(GGally)
library(ggfortify)
library(mosaic)
library(janitor)
```

```{r}
housing <- read_csv(here("data/housing_prices.csv"))

glimpse(housing)
```

Check data for NA values
```{r}
housing  %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```

<br>

2. We expect the `total_rooms` of houses to be strongly correlated with `total_bedrooms`. Use `ggpairs()` to investigate correlations between these two variables.

<br>
**Ans**
```{r}
ggpairs(select(housing, total_rooms, total_bedrooms))
```
Variables have a strong positive correlation (r = 0.934)


<br> 

3. So, we do find significant correlations. Let's drop `total_bedrooms` from the dataset, and use only `total_rooms` going forward.

<br>
**Ans**

```{r}
housing_tidy <- housing  %>%
  select(-total_bedrooms)
```


<br>

4. We are interested in developing a regression model for the `median_house_value` of a house in terms of the possible predictor variables in the dataset. 

  i. Use `ggpairs()` to investigate correlations between `median_house_value` and the predictors (this may take a while to run, don't worry, make coffee or something).

<br>
**Ans**

```{r}
# ggpairs(housing_tidy, columns = 3:9)
ggpairs(housing_tidy, columns = 1:9)
```
``median_house_value`` has a strong positive correlation with ``median_income``. This is followed by very weak correlations with ``latitude`` (negative) and ``total_rooms`` (positive). The histograms with ``ocean_proximity`` show some variation.

  ii. Perform further `ggplot` visualisations of any significant correlations you find.

<br>
**Ans**

```{r}
housing_tidy %>%
  ggplot(aes(x = median_income, y = median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
housing_tidy %>%
  ggplot(aes(x = latitude, y = median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
housing_tidy %>%
  ggplot(aes(x = total_rooms, y = median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
housing_tidy %>%
  ggplot(aes(x = ocean_proximity, y = median_house_value)) +
  geom_boxplot()
```
<br>

5. Shortly we may try a regression model to fit the categorical predictor `ocean_proximity`. Investigate the level of `ocean_proximity` predictors. How many dummy variables do you expect to get from it?

<br>
**Ans**
```{r}
housing %>% 
  distinct(ocean_proximity)
```

We expect to get 4 (= 5-1) dummy variables


<br>
 
6. Start with simple linear regression. Regress `median_house_value` on `median_income` and check the regression diagnostics.

<br>
**Ans**


Regression and Diagnostics:
```{r}
model_1 <- lm(median_house_value ~ median_income, data = housing_tidy)

autoplot(model_1)

summary(model_1)
```
The model is not adequate because:
1. The Residuals vs Fitted plot are not randomly scattered around zero.
2.The residuals are not following a normal distribution. The extreme values in the Q-Q plot are not lying on the straight line.
3. The Scale-Location plot shows that the residual variation is not occurring on a fixed band above the x-axis. The scatter shows a funnel shaped distribution, which indicates heteroskedasticity.  


<br> 

7. Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.

<br>
**Ans**
Add ``ocean_proximity`` into the model:
```{r}
model_2 <- lm(median_house_value ~ median_income + ocean_proximity, data = housing_tidy)

autoplot(model_2)

summary(model_2)
```

Diagnostics Review:

The model is not adequate because:
1. The Residuals vs Fitted plot are not randomly scattered around zero.
2.The residuals are not following a normal distribution. The extreme values in the Q-Q plot are not lying on the straight line, although  improved when compared to ``model_1``.
3. The Scale-Location plot shows that the residual variation is not occurring on a fixed band above the x-axis. The scatter shows a funnel shaped distribution, which indicates heteroskedasticity although improved when compared to ``model_1``.  

Model Interpretation:

1. The selected predictors (``median_income`` and ``ocean_proximity``) explain 55.29% of the ``median_house_value`` variance. This is a mild improvement when compared to ``model_1``.
2. For every unit increase in the ``median_income``, the ``median_house_value`` increases by 35,118.
3. The ``median_house_value`` changes depending on the house location. If a house is on an island, the price increases by 200,480.2 if the ``median_income`` predictor remains the same.

# Extension
    
<br>

8. Try adding an interaction between `log(medium_income)` and your chosen categorical predictor. Do you think this interaction term is statistically justified?

<br>
**Ans**
Chosen categorical predictor: ocean_proximity
```{r}
model_3 <- lm(log(median_house_value) ~ log(median_income) + ocean_proximity +
                log(median_income):ocean_proximity, data = housing_tidy)
```

```{r}
autoplot(model_3)

summary(model_3)
```

There's a small improvement as the current predictors explain 60.68% of the ``log(median_house_value)`` variance. Some of the predictors related to ocean proximity are statistically not significant though. 


<br>

9. Find and plot an appropriate visualisation to show the effect of this interaction

```{r}
housing_tidy %>%
  ggplot(aes(x = log(median_income),
             y = log(median_house_value), 
             colour = ocean_proximity)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ ocean_proximity)
```

