---
title: "R Notebook"
output: html_notebook
---

<hr>

# MVP

<br>

1. Load the `housing_prices.csv` data set and undertake an initial exploration of the data. You will find details on the data set on the relevant [Kaggle page](https://www.kaggle.com/camnugent/california-housing-prices)

<br>
**Ans**

Call libraries
```{r}
library(tidyverse)
library(here)
library(GGally)
library(ggfortify)
library(mosaic)
library(janitor)
```

```{r}
housing <- read_csv(here("data/housing_prices.csv"))

glimpse(housing)
```

Check data for NA values
```{r}
housing  %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```

<br>

2. We expect the `total_rooms` of houses to be strongly correlated with `total_bedrooms`. Use `ggpairs()` to investigate correlations between these two variables.

<br>
**Ans**
```{r}
ggpairs(select(housing, total_rooms, total_bedrooms))
```
Variables have a strong positive correlation (r = 0.934)


<br> 

3. So, we do find significant correlations. Let's drop `total_bedrooms` from the dataset, and use only `total_rooms` going forward.

<br>
**Ans**

```{r}
housing <- housing  %>%
  select(-total_bedrooms)
```


<br>

4. We are interested in developing a regression model for the `median_house_value` of a house in terms of the possible predictor variables in the dataset. 

  i. Use `ggpairs()` to investigate correlations between `median_house_value` and the predictors (this may take a while to run, don't worry, make coffee or something).

  ii. Perform further `ggplot` visualisations of any significant correlations you find.

<br>
**Ans**


```{r}
# ggpairs(housing, columns = 3:9)
ggpairs(housing, columns = 1:9)
```


<br>

5. Shortly we may try a regression model to fit the categorical predictor `ocean_proximity`. Investigate the level of `ocean_proximity` predictors. How many dummy variables do you expect to get from it?

<br>
**Ans**
```{r}
housing %>% 
  distinct(ocean_proximity)
```

We expect to get 4 (= 5-1) dummy variables


<br>
 
6. Start with simple linear regression. Regress `median_house_value` on `median_income` and check the regression diagnostics.

<br>
**Ans**

Regression:
```{r}
housing %>%
  ggplot(aes(x = median_income, y = median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

Diagnostics
```{r}
model <- lm(median_house_value ~ median_income, data = housing)

autoplot(model)

summary(model)
```
The model is not adequate - the residuals are not following a normal distribution.



The histogram of the `median_house_value`` shows that the data are right skewed - perhaps good to use a log-scale to model the data.
```{r}
ggplot(housing, aes(median_house_value)) +
  geom_histogram(col = "white")
```



<br> 

7. Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.

<br>
**Ans**
Add housing median age into the model:
```{r}
model2 <- lm(median_house_value ~ median_income + housing_median_age, data = housing)

autoplot(model2)

summary(model2)
```




# Extension
    
<br>

8. Try adding an interaction between `log(medium_income)` and your chosen categorical predictor. Do you think this interaction term is statistically justified?

<br>
**Ans**
Chosen categorical predictor: ocean_proximity
```{r}
housing_dummy <- housing %>%
  fastDummies::dummy_cols(select_columns = "ocean_proximity", remove_first_dummy = TRUE, remove_selected_columns = TRUE) 

housing_dummy <- housing_dummy %>% 
  clean_names()
```

```{r}
model3 <- lm(median_house_value ~ median_income + housing_median_age + 
             # log10(median_income)*ocean_proximity_inland + 
             # log10(median_income)*ocean_proximity_island + 
             # log10(median_income)*ocean_proximity_near_bay + 
             log10(median_income)*ocean_proximity_near_ocean,
             data = housing_dummy)
plotModel(model3)


```

```{r}
autoplot(model3)

summary(model3)
```


<br>

9. Find and plot an appropriate visualisation to show the effect of this interaction

```{r}
coplot(median_house_value ~ median_income | ocean_proximity_near_ocean,
       data = housing_dummy)
```

