---
title: "R Notebook"
output: html_notebook
---


Load libraries
```{r}
library(tidyverse)
library(janitor)
library(CodeClanData)
```



View data
```{r}
game_sales
```

Short description of tibble
```{r}
str(game_sales)
```

```{r}
glimpse(game_sales)
```


Check data  for NA values
```{r}
game_sales %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```


Descriptive summaries for numeric variables
```{r}
game_sales %>% 
  select_if(is.numeric) %>% 
  select(-year_of_release) %>% 
  # summary()
  psych::describe(fast = TRUE)
```

Check how many discrete game names exist
```{r}
game_sales %>% 
  distinct(name) %>% 
  nrow()
```
Check how many discrete genres exist
```{r}
game_sales %>% 
  distinct(genre)
```


Check how many discrete years of release exist
```{r}
game_sales %>% 
  distinct(year_of_release) %>% 
  arrange(desc(year_of_release))
```

Check how many discrete publishers exist
```{r}
game_sales %>% 
  distinct(publisher)
```


Check how many discrete developers exist
```{r}
game_sales %>% 
  distinct(developer)
```

Check how many discrete ratings exist
```{r}
game_sales %>% 
  distinct(rating)
```


Check how many discrete platforms exist
```{r}
game_sales %>% 
  distinct(platform)
```


Top rated games by critics
```{r}
game_sales %>% 
  select(name, genre, publisher, platform, critic_score) %>% 
  slice_max(critic_score, n = 5)
```

```{r}
game_sales %>%
  filter(year_of_release == 2002) %>% 
  select(name, critic_score, publisher) %>% 
  slice_max(critic_score, n = 5) %>% 
  ggplot()+
  aes(x = reorder(name, critic_score), y = critic_score, fill = publisher)+
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()
```


Top rated games by users
```{r}
game_sales %>% 
  select(name, genre, publisher, platform, user_score) %>% 
  slice_max(user_score, n = 5)
```

```{r}
game_sales %>%
  filter(year_of_release == 2008) %>% 
  select(name, user_score, publisher) %>% 
  slice_max(user_score, n = 5) %>% 
  ggplot()+
  aes(x = reorder(name, user_score), y = user_score, fill = publisher)+
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()
```



Top selling games
```{r}
game_sales %>% 
  select(name, genre, publisher, developer, sales, critic_score, user_score) %>% 
  slice_max(sales, n = 10)
```


```{r}
game_sales %>% 
  filter(year_of_release == 2002) %>% 
  select(name, sales, genre) %>% 
  slice_max(sales, n = 5) %>% 
  ggplot()+
  aes(x = reorder(name, sales), y = sales, fill = genre)+
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()
```

Which platform has the most games?
```{r}
game_sales %>% 
  group_by(platform) %>% 
  summarise(game_number = n()) %>% 
  arrange(desc(game_number))
```

```{r}
game_sales %>% 
  filter(year_of_release == 2002) %>% 
  group_by(platform) %>% 
  summarise(game_number = n()) %>% 
  slice_max(game_number, n = 5) %>% 
  ggplot()+
  aes(x = reorder(platform, game_number), y = game_number)+
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()
```


Which publisher has published the most games?
```{r}
game_sales %>% 
  group_by(publisher) %>% 
  summarise(game_number = n()) %>% 
  arrange(desc(game_number))
```

```{r}
game_sales %>% 
  filter(year_of_release == 2015) %>% 
  group_by(publisher) %>% 
  summarise(game_number = n()) %>% 
  slice_max(game_number, n = 5) %>% 
  ggplot()+
  aes(x = reorder(publisher, game_number), y = game_number)+
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()
```


Which developer has developed the most games?
```{r}
game_sales %>% 
  group_by(developer) %>% 
  summarise(game_number = n()) %>% 
  arrange(desc(game_number))
```

```{r}
game_sales %>% 
  group_by(developer) %>% 
  filter(year_of_release == 2015) %>%
  summarise(game_number = n()) %>% 
  slice_max(game_number, n = 5) %>% 
  ggplot()+
  aes(x = reorder(developer, game_number), y = game_number)+
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()
```



Which publisher has the highest critic score rating on average?
```{r}
game_sales %>% 
  group_by(publisher) %>% 
  summarise(avg_crit_rating = mean(critic_score), avg_user_rating = mean(user_score)) %>%
  arrange(desc(avg_crit_rating))
```
```{r}
game_sales %>% 
  filter(year_of_release == 2000) %>%
  group_by(publisher) %>% 
  summarise(avg_crit_rating = mean(critic_score), avg_user_rating = mean(user_score)) %>%
  slice_max(avg_crit_rating, n = 5) %>% 
  ggplot()+
  aes(x = reorder(publisher, avg_crit_rating), y = avg_crit_rating)+
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()
```




Which publisher has the highest user score rating on average?
```{r}
game_sales %>% 
  group_by(publisher) %>% 
  summarise(avg_crit_rating = mean(critic_score), avg_user_rating = mean(user_score)) %>%
  arrange(desc(avg_user_rating))
```

```{r}
game_sales %>% 
  filter(year_of_release == 2000) %>%
  group_by(publisher) %>% 
  summarise(avg_crit_rating = mean(critic_score), avg_user_rating = mean(user_score)) %>%
  slice_max(avg_user_rating, n = 5) %>% 
  ggplot()+
  aes(x = reorder(publisher, avg_user_rating), y = avg_user_rating)+
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip()
```


Which developer has the highest critic score rating on average?
```{r}
game_sales %>% 
  filter(year_of_release == 2000) %>%
  group_by(developer) %>% 
  summarise(avg_crit_rating = mean(critic_score), avg_user_rating = mean(user_score)) %>%
  arrange(desc(avg_crit_rating))
```


Which developer has the highest user score rating on average?
```{r}
game_sales %>% 
  filter(year_of_release == 2000) %>%
  group_by(developer) %>% 
  summarise(avg_crit_rating = mean(critic_score), avg_user_rating = mean(user_score)) %>%
  arrange(desc(avg_user_rating))
```

View number games released each year
```{r}
game_sales %>%
  group_by(year_of_release) %>% 
  summarise(game_number = n(), sales_number = sum(sales))
```
Plot histogram of game releases
```{r}
game_sales %>%
  ggplot() +
  aes(x = year_of_release) +
  geom_bar()
```

Game sales per year
```{r}
game_sales %>%
  group_by(year_of_release, genre) %>% 
  summarise(sales_number = sum(sales))
```

Plot histogram of game sales
```{r}
game_sales %>%
  group_by(year_of_release) %>% 
  summarise(sales_number = sum(sales)) %>% 
  ggplot() +
  aes(x = year_of_release, y = sales_number) +
  geom_bar(stat = "identity")
```




Check correlation btw critic score and user score
```{r}
game_sales %>%
  ggplot(aes(x = user_score, y = critic_score)) +
  geom_point() +
  geom_smooth(method = "lm", show.legend = FALSE) +
  xlim(0, 10) +
  ylim(0, 100)
```


Correlation between releases and sales 
```{r}
game_sales %>%
  group_by(year_of_release) %>% 
  summarise(game_number = n(), sales_number = sum(sales)) %>% 
  ggplot(aes(x = game_number, y = sales_number)) +
  geom_point()  +
  geom_text(
    aes(label = year_of_release),
    size = 3,
    nudge_x = 6,
    nudge_y = 1,
  ) +
  geom_smooth(method = "lm", show.legend = FALSE)+
  xlim(0, 200) +
  ylim(0, 250)
  
```







