---
title: "R Notebook"
output: html_notebook
---

# Homework Info

## Learning Objectives<br>

* Practice making plots with ggplot2
* Create your own theme and use it with plots

## Background

QikBit is a start-up company producing low cost fitness trackers. Their flagship product, the KwikBit Unit1 (spot the mistake on the patent form) reproduces the functionality of the best-in-class commercial fitness trackers, but is cheap to make.  They are currently preparing to pitch to a potential investor and require a selection of charts to include in their presentation and promotional materials.

## Data

The data for this assignment is provided in the `CodeClanData` package.

* `qb_revenue_breakdown`
* `qb_monthly_sales`
* `qb_competitors`
* `qb_device_data`

Take a little time before you start to familiarise yourself with the data.

```{r}
library(tidyverse)
library(janitor)
library(CodeClanData)
```


```{r}
qb_revenue_breakdown
qb_monthly_sales
qb_competitors
qb_device_data
```



# MVP

## Part 1

**Question 1.**  
Take the data in the dataset `qb_revenue_breakdown` and make a stacked bar chart showing the sources of revenue across the two years in the dataset.
<br>

**Ans 1.**
```{r}
qb_revenue_breakdown %>%
  ggplot() +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_bar(stat = "identity")+
  scale_fill_brewer(palette = "Paired")
  
```



**Question 2.**  
Make a line chart showing monthly sales of the "flagship product" - the 'KwikBit Unit 1' - over the last year. This data is inside `qb_monthly_sales`, but you will need to use `filter()` or `subset()` (base `R)`.
<br>

**Ans 2.**
```{r}
qb_monthly_sales %>%
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  aes(x = Date, y = Amount) +
  geom_line() +
  labs(x = "Date", y = "KwikBit Unit 1 Monthly Sales") + 
  geom_point()
```


<br>
**Question 3.**  
Make a line chart showing monthly revenue and costs over the last year. This data is also in `qb_monthly_sales`.								
<br>

**Ans 3.**
```{r}
qb_monthly_sales %>%
  filter(Cashflow == "Total Revenue" | Cashflow == "Cost") %>% 
  ggplot() +
  aes(x = Date, y = Amount, colour = Cashflow) +
  geom_line() +
  geom_point()
```


<br>
**Question 4.**  
Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use `geom_area`). This data is in `qb_competitors`.
<br>

**Ans 4.**
```{r}
qb_competitors %>% 
  ggplot() +
  aes(x = Year, y = Revenue, fill = Company) +
  geom_area() +
  scale_fill_brewer(palette = "Dark2")
```


<br>
**Question 5.**  
Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.
<br>

**Ans 5.**
```{r}
qb_competitors %>%
  ggplot() +
  aes(x = Year, y = Revenue, colour = Company) +
  geom_line() +
  geom_point() +
  scale_colour_brewer(palette = "Dark2")
```


<br>
**Question 6.**  
Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in `qb_device_data`.
<br>

**Ans 6.**
```{r}
qb_device_data %>%
  mutate(hrs_decimal = hours + mins/60) %>%
  ggplot() +
  aes(x = hrs_decimal, y = counts, group = id) +
  geom_line() +
  facet_grid(id ~ device)
```


<br>

## Part 2 

**Question 1.**  
Take the plots that you produced in part one and now polish them by:

1. Adding appropriate labels
2. Changing the scales and coordinates when appropriate.
3. Applying a unified theme, which is described below:
  * Graphs should have white backgrounds, and use colour sparingly.
  * There should be faint grid lines.
  * Font sizes should be ~12pt although titles should be slightly larger and axis labels can be slightly smaller. 
  * All plots should use colours from the following company colour scheme.


```{r}
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")
```
<br>

**Ans**
Create new theme:
```{r}
theme_hwk_week05 <- 
  theme(
    text = element_text(size = 12),
    title = element_text(size = 14),
    axis.text = element_text(size = 10),
    panel.background = element_rect(fill = "white"),
    panel.grid = element_line(colour = "grey90", linetype = "dashed")
  )
```


<br>
Updated Ans 1:
```{r}
ggplot(qb_revenue_breakdown) +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_col() +
  scale_fill_manual(values = col_scheme) +
  ggtitle(
    "QikBit Revenue Breakdown by Product Line",
    subtitle =  "(years 2018 and 2019)"
  ) +
  theme_hwk_week05
  
```

<br>
Updated Ans 2:
```{r}
qb_monthly_sales %>%
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  aes(x = Date, y = Amount) +
  geom_line(size = 2, colour = col_scheme[2]) +
  theme_hwk_week05 +
  labs(x = "Date", y = "Sales (£)") + 
  ggtitle(
    "KwikBit Unit 1 Monthly Sales",
    subtitle =  "(Aug 2018 to Jul 2019)") 
```

<br>
Updated Ans 3:
```{r}
qb_monthly_sales %>%
  filter(Cashflow == "Total Revenue" | Cashflow == "Cost") %>% 
  ggplot() +
  aes(x = Date, y = Amount, colour = Cashflow)  +
  theme_hwk_week05 +
  geom_line(size = 2)+
  scale_colour_manual(values = col_scheme) +
  labs(x = "Date", y = "Amount (£)") + 
  ggtitle("QikBit Revenue and Costs", subtitle = "(Aug 2018 to July 2019)")
```


<br>
Updated Ans 4:
```{r}
ggplot(qb_competitors) +
  aes(x = Year, y = Revenue, fill = Company) +
  geom_area() +
  scale_y_continuous(labels = scales::dollar) +
  theme_hwk_week05 +
  scale_fill_manual(values = col_scheme) +
  ggtitle("Fitness Tracker Market Revenue by Company", subtitle = "(Aug 2018 to July 2019)")
  
```

<br>
Updated Ans 5:
```{r}
qb_competitors %>%
  ggplot() +
  aes(x = Year, y = Revenue, colour = Company) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(x = "Date", y = "Sales (£)") + 
  theme_hwk_week05 +
  scale_colour_manual(values = col_scheme) +
  ggtitle("Competitor Sales", subtitle = "(2015 to 2019)")
```



<br>
Updated Ans 6:
```{r}
qb_device_data %>%
  mutate(hrs_decimal = hours + mins/60) %>%
  ggplot() +
  aes(x = hrs_decimal, y = counts, colour = id, group = id) +
  geom_line() +
  theme_hwk_week05 +
  scale_colour_manual(values = col_scheme) +
  labs(x = "Time (hrs)", y = "Steps") + 
  ggtitle("Step Count Comparison btw KwikBit Unit1 and Competitors") +
  facet_grid(id ~ device)

```




# Extension

**Question 1.**  
Try out different ways of plotting these three datasets. Which graphs do you think are particularly effective at showing QikBit's success?
<br>