---
title: "R Notebook"
output: html_notebook
---


1. Call libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(GGally)
```



2. View data
```{r}
avocado <- read_csv(here("data/avocado.csv")) %>% 
  clean_names()

glimpse(avocado)
```

3. Check for NA values
```{r}
avocado %>% 
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```
No missing values!



4. Data familiarisation

```{r}
avocado %>% 
  distinct(year)
```
This variable is categorical - should be represented as a factor

```{r}
avocado %>% 
  distinct(type)
```
Variable ``type`` is categorical --> dummy variable!


```{r}
avocado %>% 
  distinct(region)
```
There is region value that refers to the TotalUS - perhaps better to model this value rather than all of the region values together.



```{r}
avocado %>% 
  group_by(year, type) %>% 
  summarise(avg_mean_price = mean(average_price))
```
Funny jump in avocado price in 2017 - needs further investigation?


```{r}
avocado %>% 
  filter(region == "TotalUS") %>% 
  arrange(date)
```

```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  group_by(year, type) %>% 
  summarise(avg_mean_price = mean(average_price))
```
Funny jump in avocado price in 2017. I saw this previously when considered all regions - needs further investigation?


4.1 View data as time series:

```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  ggplot() +
  geom_line(aes(x = date, y = average_price, group = type, color = type))
```
Country-wide there's a seasonality in the avocado price - it peaks around September. The average price during this time of the year has been increasing as seen in the data for 2015-2017.

There were 6 consecutive weeks in 2015 (2015-07-05	to 2015-08-09) where the ``organic`` avocado price was stuck at 1.00. This value does not look normal as seen in the trend plot above. Perhaps better to replace these values with the value from 2015-06-28	(ie 1.64) or the average of avocado prices between 2015-06-28 and 2015-08-16	(ie average of 1.64 and 1.75).

NB: For the ``conventional`` avocado price, the price between 2015-07-12 and 2015-07-26 was 1.07. It is likely that these values were imputed, too.	


4.2 Check of trends for other variables
```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  ggplot() +
  geom_line(aes(x = date, y = total_volume, group = type, color = type))
```
```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  ggplot() +
  geom_line(aes(x = date, y = x4046, group = type, color = type))
```

```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  ggplot() +
  geom_line(aes(x = date, y = x4225, group = type, color = type))
```

```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  ggplot() +
  geom_line(aes(x = date, y = x4770, group = type, color = type))
```

```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  ggplot() +
  geom_line(aes(x = date, y = total_bags, group = type, color = type))
```

```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  ggplot() +
  geom_line(aes(x = date, y = small_bags, group = type, color = type))
```
```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  ggplot() +
  geom_line(aes(x = date, y = large_bags, group = type, color = type))
```

```{r}
avocado %>% 
  filter(region == "TotalUS") %>%
  ggplot() +
  geom_line(aes(x = date, y = x_large_bags, group = type, color = type))
```

The values corresponding to ``conventional`` type avocados are many orders of magnitude higher than the ``organic`` type avocados. We have to scale the data!!!



4.3 Check of avocado price trends for other regions:
```{r}
avocado %>% 
  filter(region == "West") %>%
  ggplot() +
  geom_line(aes(x = date, y = average_price, group = type, color = type))
```

```{r}
avocado %>% 
  filter(region == "Northeast") %>%
  ggplot() +
  geom_line(aes(x = date, y = average_price, group = type, color = type))
```

```{r}
avocado %>% 
  filter(region == "NewYork") %>%
  ggplot() +
  geom_line(aes(x = date, y = average_price, group = type, color = type))
```

5. Modeling Objective 

Model the avocado ``average_price`` for region ``TotalUS`` for available data of years 2015-2018


Stuff to do to the data set:

1. Variable ``type`` is categorical --> dummy variable! Have to make it a factor!

2. Variable ``year`` is categorical --> should be represented as a factor

3. ``total_bags`` equals ``small_bags`` + ``large_bags`` + ``x_large_bags``. We can drop ``total_bags``

4. The values corresponding to ``conventional`` type avocados are many orders of magnitude higher than the ``organic`` type avocados. We have to scale data!!!

5. Country-wide there's a seasonality in the avocado price - it peaks around September. The average price during this time of the year has been increasing as seen in the data for 2015-2017.

There were 6 consecutive weeks in 2015 (2015-07-05	to 2015-08-09) where the ``organic`` avocado price was stuck at 1.00. This value does not look normal as seen in the trend plot above. Perhaps better to replace these values with the value from 2015-06-28	(ie 1.64) or the average of avocado prices between 2015-06-28 and 2015-08-16	(ie average of 1.64 and 1.75).

NB: For the ``conventional`` avocado price, the price between 2015-07-12 and 2015-07-26 was 1.07. It is likely that these values were imputed, too.	



5.1 Tidy data set

```{r}
impute_value <- avocado$average_price[avocado$date == "2015-06-28" & avocado$region == "TotalUS" & avocado$type == "organic"]

avocado_tidy <- avocado %>%
  filter(region == "TotalUS") %>%
  mutate(year = as.factor(year),
         average_price = ifelse(
           type == "organic" & average_price == 1, 
           impute_value, 
           average_price),
  ) %>% 
  mutate(across(.cols = total_volume:x_large_bags, .fns = ~scale(.x))) %>% 
  select(-c(x1, date, region, total_bags))

```
5.2 Check for aliases

```{r}
alias(average_price ~ ., data = avocado_tidy)
```
Ok, no aliases occurring to data set!



5.3 Check associations of potential predictors

Split data in numeric and non-numeric
```{r}
avocado_tidy_numeric <- avocado_tidy %>%
  select_if(is.numeric)

avocado_tidy_nonnumeric <- avocado_tidy %>%
  select_if(function(x) !is.numeric(x))

avocado_tidy_nonnumeric$average_price <- avocado_tidy$average_price
```




```{r}
avocado_tidy_numeric %>%
  ggscatmat()
```




```{r}
avocado_tidy_nonnumeric %>%
  ggpairs()
```


Automated Modeling
==================



```{r}
library(leaps)
```
1. Forward search
```{r}
regsubsets_forward <- regsubsets(average_price ~ ., data = avocado_tidy, nvmax = 9, method = "forward")
```

```{r}
sum_regsubsets_forward <- summary(regsubsets_forward)
sum_regsubsets_forward
```


```{r}
plot(regsubsets_forward, scale = "bic")
```


```{r}
plot(sum_regsubsets_forward$rsq, type = "b")
```


```{r}
plot(sum_regsubsets_forward$bic, type = "b")
```

```{r}
plot(sum_regsubsets_forward$cp, type = "b")
```

Go ahead and extract the 3-predictor model, and then check its significance and the diagnostic plots.
```{r}
summary(regsubsets_forward)$which[3,]
```

```{r}
mod_with_year <- lm(average_price ~ x4225 + year, data = avocado_tidy)
summary(mod_with_year)
```

```{r}
par(mfrow = c(2, 2))
plot(mod_with_year)
```

2. Exhaustive search
```{r}
regsubsets_exhaustive <- regsubsets(average_price ~ ., data = avocado_tidy, nvmax = 9, method = "exhaustive")
```

```{r}
sum_regsubsets_exhaustive <- summary(regsubsets_exhaustive)
sum_regsubsets_exhaustive
```

```{r}
plot(sum_regsubsets_exhaustive$rsq, type = "b")
```

```{r}
plot(sum_regsubsets_exhaustive$bic, type = "b")
```


Go ahead and extract the 4-predictor model, and then check its significance and the diagnostic plots.
```{r}
summary(regsubsets_exhaustive)$which[4,]
```

```{r}
mod_exhaustive_with_year <- lm(average_price ~ total_volume + x4770 + x_large_bags + type + year, data = avocado_tidy)
summary(mod_exhaustive_with_year)
```

```{r}
par(mfrow = c(2, 2))
plot(mod_exhaustive_with_year)
```

3. Use the ``gmulti`` package

```{r}
library(glmulti)
glmulti_fit <- glmulti(
  average_price ~ ., 
  data = avocado_tidy,
  level = 2, # 2 = include pairwise interactions, 1 = main effects only (main effect = no pairwise interactions)
  minsize = 0, # no min size of model
  maxsize = -1, # -1 = no max size of model
  marginality = TRUE, # marginality here means the same as 'strongly hierarchical' interactions, i.e. include pairwise interactions only if both predictors present in the model as main effects.
  method = "g", # the problem is too large for exhaustive search, so search using a genetic algorithm
  crit = bic, # criteria for model selection is BIC value (lower is better)
  plotty = FALSE, # don't plot models as function runs
  report = TRUE, # do produce reports as function runs
  confsetsize = 100, # return best 100 solutions
  fitfunction = lm # fit using the `lm` function
)
```


```{r}
summary(glmulti_fit)
```


```{r}
# here is the best model (lowest BIC) from the glmulti run above
mod_interact <- lm(average_price ~ 1 + year + total_volume + x4046 + x4225 + x4770 +
                     small_bags + large_bags + type + small_bags:total_volume +
                     small_bags:x4225 + large_bags:total_volume + large_bags:x4046 +
                     type:total_volume + type:x4225 + type:x4770 + type:small_bags +
                     year:small_bags + year:type,
                     data = avocado_tidy)
summary(mod_interact)
```



```{r}
# let's see diagnostic plots
par(mfrow = c(2, 2))
plot(mod_interact)
```



```{r}
# calculate predictor importances. This will take a while to run, as lmg tries all possible orders of including predictors!
# library(relaimpo)
# calc.relimp(mod_interact, type = "lmg", rela = TRUE)

```







