---
title: "Probability Homework"
output: html_notebook
---


# MVP

## The Titanic


A data set detailing the survival or death of passengers and crew members on the RMS Titanic is given in the `data` folder for this homework - use this data and **not** the built-in `Titanic` dataset in `R`. 

Some time after the disaster a journalist wrote that crew members were *more likely* to have survived the sinking than passengers. This was disputed by a descendant of one of the surviving crew members, who suggested that the opposite was true, i.e. passengers were *more likely* to have survived than crew.  

**Question**  
Calculate the conditional probabilities $\textrm{prob}(\textrm{survived} \mid \textrm{crew})$ and $\textrm{prob}(\textrm{survived} \mid \textrm{passenger})$ to settle this dispute.

Have a look at the distinct values of the `class` variable, this will tell you whether each individual was a passenger (1st, 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew)

**Loading packages**

You may want to use `prob` and `gtools` later in the homework, and the `tidyverse` for this section, but if you load `prob` after `tidyverse`, `dplyr::filter()` will be masked. So load your packages in this order:

```{r, message = FALSE}
library(prob)
library(gtools)
library(tidyverse)
```

<details>
<summary>**Hints**</summary>

* You may want to create a `crew` column containing values `TRUE` or `FALSE` to help with this
* For $\textrm{prob}(\textrm{survived} \mid \textrm{crew})$ think 'what proportion of the crew survived?', and similar for $\textrm{prob}(\textrm{survived} \mid \textrm{passenger})$ 
</details>
<br>

<br>


<br>
**Ans **

Load libraries:
```{r}
library(prob)
library(gtools)
library(tidyverse)
```

View data:
```{r}
titanic <- read_csv("data/titanic.csv")

glimpse(titanic)
```

Pull classes:
```{r}
titanic %>% 
  distinct(class) %>% 
  pull()
```

Pull survival rating:
```{r}
titanic %>% 
  distinct(survived) %>% 
  pull()
```


Calculate prob(survived|crew) and prob(survived|passenger) 
```{r}
#prob(survived|crew) calculation  
prob_surv_crew <- titanic %>% 
  mutate(crew = if_else(class %in% c("engineering crew",
                                     "victualling crew", 
                                     "restaurant staff",
                                     "deck crew"), 
                        TRUE, 
                        FALSE)) %>% 
  filter(survived == "yes" & crew == TRUE) %>% 
  summarise(prob_surv_crew = n() / nrow(titanic))


prob_crew <- titanic %>% 
  mutate(crew = if_else(class %in% c("engineering crew",
                                     "victualling crew", 
                                     "restaurant staff",
                                     "deck crew"), 
                        TRUE, 
                        FALSE)) %>% 
  filter(crew == TRUE) %>% 
  summarise(prob_crew = n() / nrow(titanic))


prob_surv_given_crew <- prob_surv_crew / prob_crew 

prob_surv_given_crew


#prob(survived|passenger) calculation
prob_surv_psgr <- titanic %>% 
  mutate(crew = if_else(class %in% c("engineering crew",
                                     "victualling crew", 
                                     "restaurant staff",
                                     "deck crew"), 
                        TRUE, 
                        FALSE)) %>% 
  filter(survived == "yes" & crew == FALSE) %>% 
  summarise(prob_surv_passenger = n() / nrow(titanic))


prob_psgr <- titanic %>% 
  mutate(crew = if_else(class %in% c("engineering crew",
                                     "victualling crew", 
                                     "restaurant staff",
                                     "deck crew"), 
                        TRUE, 
                        FALSE)) %>% 
  filter(crew == FALSE) %>% 
  summarise(prob_psgr = n() / nrow(titanic))


prob_surv_given_psgr <- prob_surv_psgr / prob_psgr 

prob_surv_given_psgr

```

Conclusion: Passengers in the Titanic had 1.6 times more chances to have survived than the crew.

<br>

## Probability Consolidation Questions

Today was pretty full-on! Here are some problems to help you consolidate what you learned. Feel free to work on paper or in `R`, whichever you are most comfortable with:

**MVP - try _four_ of the problems**  
**Extension - try more than four!**

<br>
**Question 1.**  
If I roll a six-sided dice once, what is the probability of getting a number less than 5?
<br>
**Ans 1.** 
```{r}
s_one_die <- rolldie(1)

s_one_die %>% 
  dplyr::filter(X1 < 5) %>% 
  summarise(prob = n() / nrow(s_one_die))

```


<br>
**Question 2.**  
How many outcomes are there in the sample space of an experiment in which we toss three dice?
<br>
**Ans 2.** 
```{r}
s_three_dice <- rolldie(3)

s_three_dice %>% 
  summarise(size_of_sample = n())
```

<br>

**Question 3.**  
In an experiment where we roll four dice:

  * how many outcomes are in the event where the total of the four dice equals $5$, $6$ or $7$)?
  * what is the probability of this event?
  * what assumptions do we apply in answering part ii?
  
<details>
<summary>**Hint**</summary>
For part i, `rowSums(.)` may help you setup up the necessary random variable `total` (or you can use `pivot_longer()`)
</details>
<br>
**Ans 3.** 
```{r}
# sample space for roll of four dice
s_four_dice <- rolldie(4)

# event space for a roll of four dice equal to 5, 6 or 7
s_four_dice %>%
  mutate(total = rowSums(.)) %>% # equal to coding: total = X1 + X2 + X3 + X4
  dplyr::filter(total %in% c(5, 6, 7)) %>%
  summarise(size_of_event_space = n())

# probability for a roll of four dice equal to 5, 6 or 7
s_four_dice %>%
  mutate(total = rowSums(.)) %>%
  dplyr::filter(total %in% c(5, 6, 7)) %>% 
  summarise(prob = n() / nrow(s_four_dice))

# The above probability assumes that the dice are unbiased, that is the outcomes of the dice rolls are equally likely
```



<br>

**Question 4.**  
Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

<details>
<summary>**Hint**</summary>
This is an arrangement where order matters. Of all the possible arrangements, how many are correct?
</details>
<br>
**Ans 4.** 
```{r}
# create vector of increasing heights - contains 5 people, n = 5. 
heights <- c(1:5)

# Choose all five people at a time, r = 5
# Don't replace a person once chosen: repeats.allowed = FALSE
chairs <- as_tibble(permutations(n = 5, r = 5, v = heights, repeats.allowed = FALSE))

# rename the variables in chairs to something better
chairs <- chairs %>%
  rename(
    chair1 = V1,
    chair2 = V2,
    chair3 = V3,
    chair4 = V4,
    chair5 = V5
  )

# let's see what the chairs matrix looks like
chairs

# Get the event space (filter) and then calculate the prob (summarise)
chairs %>%
  filter(chair1 == 5, chair2 == 4, chair3 == 3, chair4 == 2, chair5 == 1) %>% 
  summarise(prob = n() / nrow(chairs))
```


<br>
**Question 5.**  
I roll six dice:

  * what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?
  * what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?
  
<details>
<summary>**Hint**</summary>
Have a look at the `isin()` function in the `prob` package. For example, to `filter` all rows containing $6$, $1$, $2$ in that order, you could use `isin(., c(6, 1, 2), ordered = TRUE)`
</details>
<br>
**Ans 5.** 
```{r}
# create vector of increasing die numbers - contains 6 dice, n = 6. 
dice_event <- c(1:6)

# Roll all six dice at a time, r = 6
# repeats.allowed = TRUE
dice <- as_tibble(permutations(n = 6, r = 6, v = dice_event, repeats.allowed = TRUE))

# rename the variables in dice to something better
dice <- dice %>%
  rename(
    die1 = V1,
    die2 = V2,
    die3 = V3,
    die4 = V4,
    die5 = V5,
    die6 = V6
  )

# let's see what the dice matrix looks like
dice

# Get the event space (filter) and then calculate the prob (summarise) of seeing
# the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right
dice %>%
  filter(isin(., c(6:1), ordered = TRUE)) %>% 
  summarise(prob_ordered = n() / nrow(dice))

# Get the event space (filter) and then calculate the prob (summarise) of seeing
# each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order
dice %>%
  filter(isin(., c(6:1), ordered = FALSE)) %>% 
  summarise(prob_order_not_matter = n() / nrow(dice))
```

<br>
**Question 6.**  
If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades? <br>

<details>
<summary>**Hints**</summary>

* Investigate the `cards()` function in the `prob` package, if you'd like to try this in R (make a deck without jokers). 
* You could think about adding your own `black_card` random variable to the sample space, which would be `TRUE` if the `suit` is 'Spade' or 'Club', and `FALSE` otherwise.
</details>
<br>
**Ans 6.** 
-
```{r}
deck <- cards(jokers = FALSE, makespace = FALSE)

prob_black <- deck%>% 
  mutate(black_card = suit == "Spade" | suit == "Club") %>% 
  filter(black_card == TRUE) %>% 
  summarise(prob_black = sum(black_card) / nrow(deck)) %>% 
  pull()

prob_spade_and_black <- deck%>% 
  mutate(black_card = suit == "Spade" | suit == "Club") %>% 
  filter(suit == "Spade" & black_card == TRUE) %>% 
  summarise(prob_spade_and_black = n() / nrow(deck))%>% 
  pull()


prob_spade_given_black <- prob_spade_and_black / prob_black

prob_spade_given_black
```

<br>

# Extension

Answer more than 4 consolidation questions

**Ans. ** 
Done!