---
title: "Homework - Confidence Intervals"
output: html_notebook
---


<hr>

# MVP

Now we'll go back to CI creation in the normal fashion. We'll take the `ames` data from the CIs lab earlier today and regard it now as a **sample**, we won't be drawing any smaller samples from within it. This is the usual situation in an analysis: you use all the data available to you!

<br>

**Task 1.**  
Load the data again, `clean_names()`, and re-familiarise yourself with it

<br>

**Ans 1. **

Load libraries:
```{r}
library(tidyverse)
library(infer)
library(janitor)
```

Load and explore the data held in ames.csv
```{r}
ames <- read_csv("data/ames.csv") %>%
  clean_names()
```

```{r}
glimpse(ames)
```


Check data  for NA values
```{r}
ames %>%
  summarise(across(.cols = everything(),
                   .fns = ~sum(is.na(.x))))

```


<br>

**Task 2.**  
Investigate the distribution of `lot_area`. Is the distribution roughly normal? If not, what problems do you find?

<br>

**Ans 2. **

```{r}
ames %>%
  ggplot(aes(x = lot_area)) +
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.7, binwidth = 2000)
```

The distribution is not normal. There's a significant right-skew.

<br>

**Task 3.**  
Compute and visualise a bootstrap sampling distribution for the `mean(lot_area)` of the sold houses.

<br>

**Ans 3. **
```{r}
# specify that we want to look at lot_area variable, 5000 reps, and want to calculate the mean
bootstrap_sample <- ames %>%
  specify(response = lot_area) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")
```


```{r}
bootstrap_sample %>% 
  ggplot(aes(x = stat)) +
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.7)
```

<br>

**Task 4.**  
Use your bootstrap distribution to calculate a $95\%$ CI for `mean(lot_area)`, and visualise it on the distribution

<br>

**Ans 4. **
```{r}
infer_ci_95 <- bootstrap_sample %>% 
  get_confidence_interval(level = 0.95, type = "percentile")

infer_ci_95
```

```{r}
bootstrap_sample %>%
  visualise(bin = 20) +
  shade_confidence_interval(endpoints = infer_ci_95)
```


<br>

**Task 5.**  
You would like to know the `mean(lot_area)` of the sold houses with higher confidence. Calculate the $99\%$ CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the $95\%$ CI? Does that make sense?

<br>

**Ans 5. **
```{r}
infer_ci_99 <- bootstrap_sample %>% 
  get_confidence_interval(level = 0.99, type = "percentile")

infer_ci_99
```

The $99\%$ CI is wider than the $95\%$ CI. The bounds have moved away from the mean lot area.

<br>

**Task 6.**  
Calculate the point estimate of the `mean(lot_area)`

<br>

**Ans 6. **

```{r}
bootstrap_sample %>%
  summarise(mean_lot_area = mean(stat)) 
```

<hr>

# Extension

<br>

**Task 1.**  
Calculate a point estimate and $95\%$ CI for the proportion of houses in the data built before 1920. Does the number of `reps` you use matter? [Investigate `reps` from $200$ up to $50000$, memory of your laptop permitting].
<br><br>

**Ans. **
Calculation of point estimate from original sample (in this exercise, it's the population) 
```{r}
ames %>% 
  mutate(house_built_flag = if_else(year_built < 1920, "before_1920", "after_1920")) %>%
  summarise(prop_built_before_1920 = mean(house_built_flag == "before_1920"))
```


```{r}
# specify that we want to look at house_built_flag, for 200, 500, 1000, 2000, 5000, 10000, 20000, 50000 reps, and want to calculate the mean

results_reps <- tibble()

for (resamples_num in c(200, 500, 1000, 2000, 5000, 10000, 20000, 50000)){
  
  bootstrap_sample_rep <- ames %>% 
  mutate(house_built_flag = if_else(year_built < 1920, "before_1920", "after_1920")) %>%
  specify(response = house_built_flag, success = "before_1920") %>%
  generate(reps = resamples_num, type = "bootstrap") %>%
  calculate(stat = "prop")
  
  infer_ci_95 <- bootstrap_sample_rep %>% 
  get_confidence_interval(level = 0.95, type = "percentile")
  
  # bootstrap_sample_rep %>%
  # summarise(point_est = mean(stat))
  
  # prepare new row for results, storing CI bounds
  this_result <- c(
    resamples= resamples_num, 
    point_est = mean(bootstrap_sample_rep$stat),
    lower_bound = infer_ci_95 %>% 
      pull(lower_ci), 
    upper_bound = infer_ci_95 %>%
      pull(upper_ci),
    ci_width = infer_ci_95 %>% pull(upper_ci) - infer_ci_95 %>% pull(lower_ci)
  )


  # add this_result to growing tibble of results
  results_reps <- results_reps %>%
    bind_rows(this_result)
  
  
}

results_reps
```

As the number of reps increases, the mean proportion of houses built before 1920 and the 95% CI upper/lower bounds converge for more than 2000 generated resamples. The upper/lower bound estimates converge for more than 2000 reps. 






